---
title: "Homework_1"
author: "Aishwarya Thorat - 24211250"
format: 
  pdf:
    documentclass: article
    geometry: margin=1.5in
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

## Introduction:

The `ToothGrowth` dataset is a built-in dataset in R. It examines the effect of Vitamin C on tooth growth in 60 guinea pigs. The experiment compares two supplement types\
- **orange juice (OJ)**\
- **ascorbic acid (VC)**\
administered at three dose levels: 0.5, 1, and 2 mg/day.

This dataset is commonly used for visualizing grouped data and performing basic exploratory data analysis.

## Dataset Description:

The dataset contains 3 variables:

-   **`len`**: Tooth length (numeric)
-   **`supp`**: Supplement type – Orange Juice (`OJ`) or Vitamin C (`VC`)
-   **`dose`**: Dose of vitamin C in milligrams per day (numeric: 0.5, 1, 2)

These variables allow us to compare how both the supplement type and dosage level affect tooth growth in guinea pigs.

## Summary Statistics:

```{r}
summary(ToothGrowth)
```

This summary shows the range and distribution of tooth length values for different doses and supplement types.

## Plot:

The bar plot below depicts the average tooth length for each combination of supplement type and dose level.

```{r}
#| warning: false
library(ggplot2)
library(dplyr)

# Modify supplement names and calculate group means
avg_len <- ToothGrowth %>%
  mutate(supp = recode(supp, OJ = "Orange Juice", VC = "Vitamin C")) %>%
  group_by(supp, dose) %>%
  summarise(mean_len = mean(len), .groups = "drop")

avg_len$dose <- as.factor(avg_len$dose)

ggplot(avg_len, aes(x = dose, y = mean_len, fill = supp)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.2) +
  labs(
    title = "Average Tooth Length by Supplement Type and Dose",
    x = "Dose (mg/day)",
    y = "Mean Tooth Length (mm)",
    fill = "Supplement"
  ) +
  scale_fill_manual(
    values = c("Orange Juice" = "#FFA500", "Vitamin C" = "#1E90FF")
  ) +
  theme_minimal()
```

## Explanation:

-   The bar plot visualizes how tooth growth varies by dose and supplement type.
-   Bars are grouped by supplement type for each dose level.
-   Custom colors improve clarity: orange for Orange Juice and blue for Vitamin C.

## Interpretation:

-   Tooth growth increases as the dose increases.
-   At lower doses, Orange Juice results in more growth than Vitamin C.
-   At the highest dose (2 mg/day), both supplements produce similar growth.

This indicates that dose has a stronger effect on growth than supplement type, especially at high levels.
